YouTube Kids: The Effect of YouTube on Children’s Racial Learning

### 2025-03-12 manuscript.qmd
updated qmd

---
title: "YouTube Kids: The Effect of YouTube on Children’s Racial Learning"
shorttitle: "YouTube Racial Attitudes"
author: 
- name: Reilly Amera
  corresponding: true
  email: ramera@uchicago.edu
  affiliations:
  - id: id1
    name: "University of Chicago"
    department: MA Program in the Social Sciences
floatsintext: true
numbered-lines: false
bibliography: [r-references.bib, MA_THESIS.bib]
link-citations: true
draft-date: false
lang: en
language:
  citation-last-author-separator: "and"
  citation-masked-author: "Masked Citation"
  citation-masked-date: "n.d."
  citation-masked-title: "Masked Title"
  email: "Email"
format:
  apaquarto-html: default
  apaquarto-docx: default
  apaquarto-typst: 
    keep-typ: true
    list-of-figures: true
    list-of-tables: true
    toc: true
  apaquarto-pdf:
    documentmode: man
    keep-tex: true
    list-of-figures: true
    list-of-tables: true
    fig-width: 2
    fig-height: 2
    out-width: 40%
editor: visual
---

@@

library(svglite)
library(papaja)
library(knitr)
library(broom)
library(kableExtra)
library(flextable)
library(psych)
library(tidyverse)

@@

```{r}
#| label: my-theme
#| echo: false
#| message: false
#| warning: false

@@

### Literature Review {#sec-lit-review}

@@

### Present Study {#sec-present-study}

@@

### Participants and Procedures {#sec-participants-and-procedures}

@@

### Measures {#sec-measures}

@@

1. Beliefs about groups (adapted from @halperinetal_2012_promoting)
2. Perceived homophily in behavior and appearance (adapted from @mccroskeyetal_1975_development)
3. Parasocial interaction with a YouTuber (adapted from @sungetal_2023_shes)
4. Racial regularities (created based on @rollinsetal_2022_who)
5. Friend choice (adapted from @amemiyabian_2024_why)
All measures used in adult studies have been specifically adapted for child readability.

```{r}
#| label: report-reliability-function
#| echo: false
#| message: false
#| warning: false

return_alpha <- function(data, cols) {
  # select data 
  selected_data <- data %>% select(all_of(cols))
  # compute alpha
  alpha_result <- psych::alpha(selected_data, check.keys = TRUE)
  # Extract and return the raw alpha value
  return(alpha_result$total$raw_alpha)
}
```

```{r}
#| label: reliability-df
#| echo: false
#| message: false
#| warning: false

measures <-
  list(
  "Group Malleability" = c("malleable_1", "malleable_2", "malleable_3", "malleable_4"),
  "Homophily (Appearance)" = c("homophily_appear_1", "homophily_appear_2", "homophily_appear_3", "homophily_appear_4", "homophily_appear_5"),
  "Homophily (Appearance)" = c("homophily_behave_1", "homophily_behave_2", "homophily_behave_3", "homophily_behave_4", "homophily_behave_6"),
  "Parasociality" = c("parasocial_1", "parasocial_2", "parasocial_3", "parasocial_4", "parasocial_5"),
  "Racial Regularities" = c("regularities_1", "regularities_2", "regularities_3", "regularities_4", "regularities_5")
)

reliability <- data.frame(
  Measure = names(measures),
  Alpha = sapply(measures, function(cols) return_alpha(youtube, cols))
)

reliability <- reliability %>%
  mutate(Alpha = round(Alpha, digits = 2)) %>% 
  mutate(
    Evaluation = case_when(
      Alpha >= 0.95 ~ "Excellent",
      Alpha >= 0.9 ~ "Great",
      Alpha >= 0.8 ~ "Good",
      Alpha >= 0.7 ~ "Acceptable",
      Alpha >= 0.6 ~ "Questionable",
      Alpha < 0.6 ~ "Poor"
    )
  )
```

```{r}
#| label: tbl-reliability
#| tbl-cap: "Questionnaire Reliability"
#| echo: false
#| message: false
#| warning: false

# Generate table
kable(reliability, 
      align = c("c", "c", "c"),
      caption = "Reliability (assessed through Cronbach's Alpha) of questionnaires used, within sample") %>% 
  kable_styling(position = "center", latex_options = c("striped"))

@@

@tbl-reliability reports the reliability of all questionnaires used, within the present sample.

@@

Looking first to the distributions of time spent on YouTube among  participants, just as many children use YouTube every day as (*N* = `r sum(youtube$daily == 'Every Day')`) as those who do not (*N* = `r sum(youtube$daily == 'Not Every Day')`). And, as shown in @fig-time-distribution, children who use YouTube every day appear to use YouTube for more time on an average day than children who do not. On average, children spend about `r round(mean(youtube$Ptime), digits = 0)` minutes, or `r round(mean(youtube$Ptime)/60, digits = 0)` hour, on YouTube on an average day.

```{r}
#| label: fig-time-distribution
#| fig-cap: "Distribution of watch time for kids"
#| fig-align: "center"
#| echo: false
#| message: false
#| warning: false

@@

ggplot(subset(youtube, !is.na(Ctime)), aes(x = Ctime)) + 
  geom_bar(color = "black", fill = "grey") +
  facet_grid(~ daily) +
  scale_x_discrete(labels = c("About 15 minutes a day" = "~ 15 min", "About 30 minutes a day" = "~ 30 min", "About 45 minutes a day" = "~ 45 min", "About 1 hour a day" = "~ 1 hr", "About 1 and 1/2 hours a day" = "~ 1.5 hr", "More than 2 hours a day" = "> 2 hr")) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  labs(x = "Time Spent on YouTube",
       y = "Number of Participants",
       title = "Time Spent on YouTube on an Average Day",
       subtitle = "Comparing children who use YouTube every day and children who do not") +
  my_theme()

@@

To what extent is time spent on YouTube predictive of how often children observe racial regularities?

```{r}
summary(aov(regularity_score ~ Ctime, data = youtube))
summary(aov(regularity_score ~ outgroup_freq, data = youtube))
```

```{r}
summary(aov(malleable_score ~ Ctime, data = youtube))
summary(aov(malleable_score ~ outgroup_freq, data = youtube))
```

To what extent is time spent on YouTube predictive

```{r}
#| label: regression 
#| echo: false
#| message: false
#| warning: false

# BASED ON APAQUARTO DEMO - WILL CHANGE FOR FINAL; WANTED TO TEST RENDERING

# regression model
model <- lm(malleable_score ~ Ptime, data = youtube)

summary(lm(malleable_score ~ regularity_score, data = youtube))

t.test(malleable_score ~ race, data = youtube)
t.test(regularity_score ~ race, data = youtube)

t.test(malleable_score ~ daily, data = youtube)
t.test(regularity_score ~ daily, data = youtube)
```

```{r}
#| label: regression-report-functions
#| echo: false
#| message: false
#| warning: false

# BASED ON APAQUARTO DEMO - WILL CHANGE FOR FINAL; WANTED TO TEST RENDERING AND TRY INLINE REFERENCES

# report_regression takes a lm model object as an argument
# and returns a string of format:
# "$F$(df, df.residual) = F-value, $p$ = p-value, $R^2$ = r-squared"
report_regression <- function(model) {

  # Extract coefficients and p-values
  m_glance <- glance(model)

  # Format the p-value
  p_value <- apa_p(m_glance$p.value, add_equals = TRUE)

  # Create a string with the results
  report <- paste0("$F$(", 
                   m_glance$df, 
                   ", ", m_glance$df.residual, 
                   ") = ", round(m_glance$statistic, 2), 
                   ", $p$ = ", p_value, ", $R^2$ = ", round(m_glance$r.squared,2))

  return(report)    
}

@@

Below is a simple regression displaying the relationship between time spent on YouTube and a child's racial attitudes. Despite a visual upward trend, this relationship is not statistically significant: `r report_regression(model)`.

```{r}
#| label: fig-regression-plot
#| fig-cap: "Simple Correlation Between Time Spent on YouTube and Racial Attitudes score"
#| fig-scap: "Time v. Attitudes"
#| fig-align: "center"
#| fig-width: 6
#| fig-height: 4
#| echo: false
#| message: false
#| warning: false

ggplot(youtube) +
  aes(x = regularity_score, y = malleable_score) +
  geom_point(aes(color = daily)) +
  stat_smooth(method = "lm") +
  facet_wrap(~ race) +
  scale_y_continuous(name = "Malleability Score") +
  scale_x_continuous(name = "Regularities Score") +
  my_theme()

```
# Discussion

>>>>

---
title: "YouTube Kids: The Effect of YouTube on Children’s Racial Learning"
shorttitle: "YouTube Racial Attitudes"
author: 
- name: Reilly Amera
  corresponding: true
  email: ramera@uchicago.edu
  affiliations:
  - id: id1
    name: "University of Chicago"
    department: MA Program in the Social Sciences
    address: "1155 E 60th St."
    city: Chicago
    region: IL
    country: USA
    postal-code: 60637
abstract: "One way in which children learn about group dynamics within their racial world is from patterns in their daily environments. The current study seeks to understand the degree to which YouTube, a popular form of social media, exposes children to negatively valenced racial patterns, and potentially affects their beliefs. Dyads of parents and 8-13 year-old children (N = 18) describe daily YouTube usage, and the effects of children's YouTube usage are assessed. In the present sample, there is no relationship between children's YouTube usage, racial patterns they may observe, and negative racial beliefs."
floatsintext: true
numbered-lines: false
bibliography: [r-references.bib, bibliography.bib]
link-citations: true
draft-date: false
lang: en
language:
  citation-last-author-separator: "and"
  citation-masked-author: "Masked Citation"
  citation-masked-date: "n.d."
  citation-masked-title: "Masked Title"
  email: "Email"
format:
  apaquarto-html: default
  apaquarto-docx: default
  apaquarto-typst: 
    keep-typ: true
    list-of-figures: true
    list-of-tables: true
    toc: true
  apaquarto-pdf:
    documentmode: man
    keep-tex: true
    list-of-figures: true
    list-of-tables: true
editor: visual
---

@@

# Necessary libraries
library(svglite)
library(papaja)
library(knitr)
library(broom)
#install.packages(kableExtra)
library(kableExtra)
library(flextable)
library(psych)
library(tidyverse)

# Setting global options
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE)

@@

```{r}
#| label: my-theme

@@

## Literature Review {#sec-lit-review}

@@

## Present Study {#sec-present-study}

@@

## Participants and Procedures {#sec-participants-and-procedures}

@@

## Measures {#sec-measures}

@@

1. Beliefs about groups (adapted from @halperinetal_2012_promoting)
2. Perceived homophily in behavior and appearance (adapted from @mccroskeyetal_1975_development)
3. Parasocial interaction with a YouTuber (adapted from @sungetal_2023_shes)
4. Racial regularities (created based on @rollinsetal_2022_who)
5. Friend choice (adapted from @amemiyabian_2024_why)

All measures used in adult studies have been specifically adapted for child readability. @tbl-reliability reports the reliability of all questionnaires used, within the present sample.

```{r}
#| label: report-reliability-function

# Function which returns Cronbach's alpha
return_alpha <- function(data, cols) {
  # select data 
  selected_data <- data %>% select(all_of(cols))
  # compute alpha
  alpha_result <- psych::alpha(selected_data, check.keys = TRUE)
  # Extract and return the raw alpha value
  return(alpha_result$total$raw_alpha)
}
```

```{r}
#| label: reliability-dataframe

# Creating a list of all measures, grouped by questionnaire
measures <-
  list(
  "Group Malleability" = c("malleable_1", "malleable_2", "malleable_3", "malleable_4"),
  "Homophily (Appearance)" = c("homophily_appear_1", "homophily_appear_2", "homophily_appear_3", "homophily_appear_4", "homophily_appear_5"),
  "Homophily (Appearance)" = c("homophily_behave_1", "homophily_behave_2", "homophily_behave_3", "homophily_behave_4", "homophily_behave_6"),
  "Parasociality" = c("parasocial_1", "parasocial_2", "parasocial_3", "parasocial_4", "parasocial_5"),
  "Racial Regularities" = c("regularities_1", "regularities_2", "regularities_3", "regularities_4", "regularities_5")
)

# Creating data frame with measures and corresponding alpha
reliability <- data.frame(
  Measure = names(measures),
  Alpha = sapply(measures, function(cols) return_alpha(youtube, cols))
)

# Adding column with evaluations of alpha
reliability <- reliability %>%
  mutate(Alpha = round(Alpha, digits = 2)) %>% 
  mutate(
    Evaluation = case_when(
      Alpha >= 0.95 ~ "Excellent",
      Alpha >= 0.9 ~ "Great",
      Alpha >= 0.8 ~ "Good",
      Alpha >= 0.7 ~ "Acceptable",
      Alpha >= 0.6 ~ "Questionable",
      Alpha < 0.6 ~ "Poor"
    )
  )
```

```{r}
#| label: tbl-reliability
#| tbl-cap: "Reliability of Questionnaires"

# Generate table
kable(reliability, 
      align = c("c", "c", "c"),
      # Setting caption here
      caption = "Reliability (assessed through Cronbach's Alpha) of questionnaires used, within sample") %>% 
  kable_styling(position = "center", latex_options = c("striped"))

@@

Beliefs about groups, particularly in terms of group malleability, were chosen as the primary outcome measure to assess racial attitudes. Such questions, adapted from @halperinetal_2012_promoting, address whether children have fixed or malleable views regarding outgroups (e.g. when prompted to think about groups with different skin colors than their own, children are asked the extent to which they agree with the statement, "These groups of people can't really change the way they act"). Higher scores in this domain demonstrate more fixed evaluations of outgroup members, suggesting an underlying racial essentialism, or a belief in the intrinsic nature of racial difference [@waxman_2021_racial].

Composite scores have been created for each measure. All analyses have been completed in RStudio [@R-base], primarily within the tidyverse and psych packages (@R-tidyverse; @R-psych).

@@

Looking first to the distributions of time spent on YouTube among  participants, just as many children use YouTube every day as (*N* = `r sum(youtube$daily == 'Every Day')`) as those who do not (*N* = `r sum(youtube$daily == 'Not Every Day')`). And, as shown in @fig-time-distribution, children who use YouTube every day appear to use YouTube for more time on an average day than children who do not. On average, parents report that children spend about `r round(mean(youtube$Ptime), digits = 0)` minutes, or `r round(mean(youtube$Ptime)/60, digits = 0)` hour, on YouTube on an average day. Some children, as parents report, watch as much as `r round(max(youtube$Ptime)/60, digits = 0)` hours of YouTube, while others watch only `r round(min(youtube$Ptime)/60, digits = 0)`. Children's self-report YouTube usage reflects a similar pattern, as shown in @tbl-frequency: the majority of participants report watching YouTube for 30 minutes to 1 hour each day.

```{r}
#| label: fig-time-distribution
#| fig-cap: "Distribution of participants' YouTube watch time"
#| fig-align: "center"
#| fig-width: 6
#| fig-height: 4

@@

ggplot(subset(youtube, !is.na(Ctime)), aes(x = Ctime)) + 
  geom_bar(color = "black", fill = "grey") +
  facet_grid(~ daily) +
  scale_x_discrete(labels = c("About 15 minutes a day" = "~ 15 min", "About 30 minutes a day" = "~ 30 min", "About 45 minutes a day" = "~ 45 min", "About 1 hour a day" = "~ 1 hr", "About 1 and 1/2 hours a day" = "~ 1.5 hr", "More than 2 hours a day" = "> 2 hr")) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  labs(x = "Time Spent on YouTube (Self-Reported)",
       y = "Number of Participants",
       title = "Time Spent on YouTube on an Average Day",
       subtitle = "Comparing children who use YouTube every day and children who do not") +
  my_theme()

@@

```{r}
#| label: generate-frequency-table

frequency <- round(addmargins(prop.table(table(youtube$Ctime, youtube$daily))), 2)
```

```{r}
#| label: tbl-frequency
#| tbl-cap: "YouTube Screen Time Frequency Table"

kable(frequency, 
      align = c("c", "c", "c"),
      # Setting caption here
      caption = "Proportions of watch times in children who watch YouTube every day vs those who do not") %>% 
  kable_styling(position = "center", latex_options = c("striped"))
```

```{r}
#| label: anova-outputs

# ANOVA: racial regularities score predicted by time spent on YT (categorical self-report)
aov_reg_ctime <- aov(regularity_score ~ Ctime, data = youtube)

# ANOVA: racial regularities score predicted by perceptions of how often kids see outgroup (categorical)
aov_reg_outfreq <- aov(regularity_score ~ outgroup_freq, data = youtube)

# ANOVA: racial regularities score predicted by perceptions of how often kids see outgroup (categorical)
aov_mall_outfreq <- aov(malleable_score ~ outgroup_freq, data = youtube)
```

```{r}
#| label: anova-report-function

# based on apaquarto manuscript example and chat help

# Takes an aov model object as argument
# Returns string of format "$F$(df1, df2) = F-value, $p$ = p-value"

report_anova <- function(model) {

  # make summary object
  m_summary <- summary(model)[[1]]
  # extract values
  df1 <- m_summary$Df[1]
  df2 <- m_summary$Df[2]
  F_value <- m_summary$`F value`[1]
  p_value <- m_summary$`Pr(>F)`[1]
  # Create a string with the results
  report <- paste0("$F$(", 
                   df1,  ", ", 
                   df2, ") = ", 
                   round(F_value, 2), 
                   ", $p$ = ", round(p_value, 3))
  # Return report
  return(report)    
}

@@

Preliminary one-way analysis of variance does not indicate that greater amounts of time spent on YouTube (self-report) is associated with a higher racial regularities score (`r report_anova(aov_reg_ctime)`). Similarly, thus far, one-way analysis of variance does not indicate that there is any association between ratings of frequency with which children encounter those of other racial groups and racial regularities scores, (`r report_anova(aov_reg_outfreq)`). Further, one-way analysis of variance does not indicate a significant relationship between the primary racial attitudes outcome measure, group malleability, and frequency of seeing racial outgroups on YouTube, (`r report_anova(aov_mall_outfreq)`).

```{r}
#| label: regression-models

# Regressing malleability against time spent on YouTube (reported by parents)
lm_mt <- lm(malleable_score ~ Ptime, data = youtube)

# Regressing malleability against time spent on YouTube (reported by parents) and whether or not YouTube is used daily
lm_mtd <- lm(malleable_score ~ Ptime + daily, data = youtube)

# Regressing malleability against regularity
lm_mr <- lm(malleable_score ~ regularity_score, data = youtube)
```

```{r}
#| label: regression-report-functions

# Taken from apaquarto demo

# report_regression takes a lm model object as an argument
# Returns string of format: "$F$(df, df.residual) = F-value, $p$ = p-value, $R^2$ = r-squared"

report_regression <- function(model) {

  # Extract coefficients and p-values
  m_glance <- glance(model)
  # Format the p-value
  p_value <- apa_p(m_glance$p.value, add_equals = TRUE)
  # Create a string with the results
  report <- paste0("$F$(", 
                   m_glance$df, 
                   ", ", m_glance$df.residual, 
                   ") = ", round(m_glance$statistic, 2), 
                   ", $p$ = ", p_value)
  # Return report
  return(report)    
}
```

Regressing malleability scores against parent reports of time spent on YouTube, no significant relationship is observed (`r report_regression(lm_mt)`). . This relationship remains unchanged and similarly insignificant when daily use is added as a predictor (`r report_regression(lm_mtd)`). Such relationships, as well as that of regularity score and YouTube usage, may be observed in @fig-malleable and @fig-regularity. Based on pilot sample alone, there is not enough data to conduct analyses with regard to the effect of YouTuber on racial attitudes, but the modal favorite YouTuber reported has thus far been `r mode(youtube$favorite)` (*N* = `r sum(youtube$favorite == 'MrBeast')`). Preliminary relationships between favorite YouTuber (this YouTuber in particular) as well as malleability and regularity scores may be found in @fig-malleable and @fig-regularity respectively.

```{r}
#| label: fig-malleable
#| fig-cap: "Correlations Between Time Spent On YouTube and Group Malleability Beliefs"
#| fig-scap: "Effects of time, daily use, and favorite YouTuber"
#| fig-note: "MrBeast is the modal favorite YouTuber. No other YouTuber has been listed with enough frequency to report."
#| fig-align: "center"
#| fig-width: 5
#| fig-height: 2

ggplot(data = youtube %>% 
             filter(Ptime > 10)) +
  aes(x = Ptime, y = malleable_score) +
  facet_grid(~ daily, scales = "free_x") +
  stat_smooth(method = "lm", color = "#800000") +
  geom_point(aes(shape = youtuber)) +
  scale_shape_manual(values = c("Not MrBeast" = 16, "MrBeast" = 1)) +
  labs(x = "Time Spent on YouTube (minutes)", y = "Malleability Score",
       shape = "Favorite YouTuber") +
  my_theme()
```

```{r}
#| label: fig-regularity
#| fig-cap: "Correlations Between Time Spent On YouTube and Racial Regularities"
#| fig-scap: "Effects of time, daily use, and favorite YouTuber"
#| fig-note: "MrBeast is the modal favorite YouTuber. No other YouTuber has been listed with enough frequency to report."
#| fig-align: "center"
#| fig-width: 5
#| fig-height: 2

ggplot(data = youtube %>% 
             filter(Ptime > 10)) +
  aes(x = Ptime, y = regularity_score) +
  facet_grid(~ daily, scales = "free_x") +
  stat_smooth(method = "lm", color = "#800000") +
  geom_point(aes(shape = youtuber)) +
  scale_shape_manual(values = c("Not MrBeast" = 16, "MrBeast" = 1)) +
  labs(x = "Time Spent on YouTube (minutes)", y = "Regularity Score",
       shape = "Favorite YouTuber") +
  my_theme()
```

Just as there is not yet an observed relationship between time spent on YouTube and racial regularities, so too is there a lack of relationship between regularities scores and malleability scores in the present sample (`r report_regression(lm_mr)`). The preliminary relationship between racial regularity and malleability scores within the full sample may be found in @fig-full. Preliminary effects of daily YouTube usage and participant race may be found in @fig-faceted.

```{r}
#| label: fig-full
#| fig-cap: "Correlation Between Racial Regularities Score and Malleability Score"
#| fig-scap: "Full Sample"
#| fig-note: "N = 18, only pilot data used."
#| fig-align: "center"
#| fig-width: 4
#| fig-height: 3

ggplot(youtube) +
  aes(x = regularity_score, y = malleable_score) +
  stat_smooth(method = "lm", color = "#800000", linewidth = 0.8, fill = "lightgrey") +
  geom_point() +
  labs(x = "Regularities Score", y = "Malleability Score",
       title = "A") +
  my_theme()
```

```{r}
#| label: fig-faceted
#| fig-cap: "Correlation Between Racial Regularities Score and Malleability Score"
#| fig-scap: "Effects of daily use and participant race"
#| fig-note: "Participant race coded as White/Non-White for pilot data only."
#| fig-align: "center"
#| fig-width: 4
#| fig-height: 3

ggplot(youtube) +
  aes(x = regularity_score, y = malleable_score) +
  stat_smooth(method = "lm", color = "black", linewidth = 0.8, fill = "lightgrey") +
  geom_point(aes(color = race)) +
  scale_color_manual(values = c("#155F83", "#C16622")) +
  facet_wrap(~ daily) +
  labs(x = "Regularities Score", y = "Malleability Score",
       color = "Participant Race",
       title = "B") +
  theme(legend.position = "bottom") +
   my_theme()

# Discussion {#sec-discussion}

The present study examines the relationship between children's YouTube usage and potential effects on racial attitudes. No significant effects of YouTube usage on group beliefs have been observed thus far; we thus fail to reject the null hypothesis that there is no difference in racial beliefs between children who spend a lot of time on YouTube and those who do not. Similarly, as there is no significant relationship between time spent on YouTube and children's observation of racial regularities, nor between racial regularities and group malleability beliefs, there is no evidence thus far to suggest that observation of racial regularities affects outgroup prejudices as a result of YouTube usage.
